% spi_slave verification report
% Koen Van Caekenberghe (koen.vancaekenberghe@chipdesign.be), ChipDesign B.V., 06.2026
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{courier}
\usepackage{listings}
\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
\title{SPI\_Slave Verification and RTL2GDS Flow}
\author{Koen Van Caekenberghe, ChipDesign B.V.}
\date{June 2026}
\begin{document}
\maketitle
\section{Overview}
The `spi_slave` design is a small SPI slave IP block that supports a simple 8-bit address and data transfer protocol. The SPI interface operates in mode 0 (CPOL=0, CPHA=0). The design exposes a secondary read port `Data2_out` to allow local readback of register contents.

\section{Design Summary}
The RTL implements the following functionality:
\begin{itemize}
  \item SPI command and data phases with a 1-byte command word.
  \item Read/write selection using the MSB of the command byte.
  \item Eight 8-bit internal registers (`Register0`..`Register7`).
  \item Synchronous register update on the internal clock domain.
  \item MISO output generation during SPI read transactions.
\end{itemize}

The top-level interface is:
\begin{itemize}
  \item `Clk` -- internal clock input.
  \item `iRST_N` -- active-low synchronous reset.
  \item `SCK`, `MOSI`, `SSEL` -- SPI inputs.
  \item `MISO` -- SPI output.
  \item `Add2_in`, `Data2_out` -- local read address and data output.
  \item `debug` -- internal debug status outputs.
\end{itemize}

\section{Place-and-Route Flow}
The repository includes a LibreLane-based RTL2GDS flow configured for the IHP SG13G2 PDK.

Important files for the flow:
\begin{itemize}
  \item `flow/Makefile` -- entry point for LibreLane and OpenROAD commands.
  \item `flow/config.yaml` -- LibreLane configuration file.
  \item `flow/run_flow.sh` -- environment setup and flow wrapper.
  \item `flow/ihp_pdk.env.example` -- example PDK environment template.
  \item `flow/constraint.sdc` -- timing constraint file for the `spi_slave` clock.
  \item `flow/place_route.tcl` -- fallback OpenROAD placement/routing script.
\end{itemize}

Reproducible flow commands:
\begin{lstlisting}
cd spi_slave_github_repo/flow
cp ihp_pdk.env.example ihp_pdk.env
# Edit ihp_pdk.env to point to the local IHP PDK installation
./run_flow.sh
\end{lstlisting}

The flow requires a valid local IHP PDK installation and the PDK name `ihp-sg13g2`.

\section{Reproducible Test Clone}
A dedicated clone of the public repo contents was created at `spi_slave/test_clone`. This isolated copy was used for all RTL and netlist verification runs, preserving the original repository contents.

\section{Verification Approach}
Two primary verification paths were exercised:
\begin{itemize}
  \item `tb_spi_slave.v` -- original RTL testbench used to validate the RTL model.
  \item `tb_spi_slave_compare.v` -- reusable comparison harness that can instantiate either the RTL design or the synthesized netlist when compiled with or without the `SYNTH` macro.
\end{itemize}

Simulation commands used for verification in `spi_slave/test_clone`:
\begin{lstlisting}
# RTL baseline with the comparison harness
iverilog -g2012 -o tb_compare_rtl.out tb_spi_slave_compare.v
vvp tb_compare_rtl.out

# Synthesized Yosys netlist with the comparison harness
iverilog -g2012 -DSYNTH -o tb_compare_synth.out flow/spi_slave_synth.v tb_spi_slave_compare.v
vvp tb_compare_synth.out

# RTL baseline using the original tb_spi_slave.v testbench
iverilog -g2012 -o tb_rtl_builtin.out tb_spi_slave.v
vvp tb_rtl_builtin.out

# Synthesized netlist using the original tb_spi_slave.v testbench with the netlist included
sed 's|`include "spi_slave.v"|`include "flow/spi_slave_synth.v"|' tb_spi_slave.v > tb_spi_slave_synth.v
iverilog -g2012 -o tb_synth_builtin.out tb_spi_slave_synth.v
vvp tb_synth_builtin.out
\end{lstlisting}

\section{Back-to-Back Results}
The observed simulation outputs were identical across the RTL and synthesized netlist runs:
\begin{itemize}
  \item RTL comparison harness result: `FINISH 750: addr=03 data=0d debug=1b miso=1 ssel=0`
  \item Synthesized Yosys netlist comparison result: `FINISH 750: addr=03 data=0d debug=1b miso=1 ssel=0`
\end{itemize}

The original `tb_spi_slave.v` testbench also completed successfully for both the RTL model and the structured Yosys netlist, confirming functional equivalence for the exercised SPI transaction.

\section{Flow Execution Results}
The RTL2GDS flow was executed from `spi_slave/flow` and produced a completed LibreLane/OpenROAD run under `spi_slave/flow/runs/RUN_2026-06-07_09-40-10`.

Key outputs from the flow include:
\begin{itemize}
  \item `runs/RUN_2026-06-07_09-40-10/final/gds/spi_slave.gds`
  \item `runs/RUN_2026-06-07_09-40-10/final/def/spi_slave.def`
  \item `runs/RUN_2026-06-07_09-40-10/final/metrics.json`
  \item `runs/RUN_2026-06-07_09-40-10/final/metrics.csv`
\end{itemize}

The flow completed with warnings only, including clock-tree synthesis warnings and various checker/report metrics that were not generated due to optional reporting steps. The run produced a valid final GDS and route database for the `spi_slave` design.

\section{Conclusions}
The `spi_slave` design behaves consistently across RTL and synthesized netlist simulation. The provided RTL2GDS flow has been exercised successfully and generated final physical outputs, and the cloned test tree in `spi_slave/test_clone` ensures that verification can be repeated cleanly.
\end{document}
